Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-16 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Adding Wall and Bench Experimental descriptions need additional refs

The cone calorimeter is probably the most versatile oxygen consumption method for measuring heat release rate. It was developed at NIST in the 1980s43 and is presently the most commonly used bench-scale rate of heat release apparatus.44 The cone calorimeter has been adopted as ASTM E1354, Test Method for Heat and Visible Smoke Release Rates for Materials and Products Using an Oxygen Consumption Calorimeter.19 It has also been adopted as an international standard, ISO 5660-1.45
The cone calorimeter consists of a heater, spark ignitor, sample holder, and load cell located underneath an exhaust hood. Typically, the sample is located in the open with free access of air to the combustion zone. The heater consists of a 5 kW electrical heating element inside an insulated stainless-steel conical shell. Samples can be tested in either a horizontal or vertical orientation. When tests are performed in the horizontal configuration, the specimen is positioned approximately 25 mm beneath the bottom plate of the cone heater. Flames and products of combustion pass through a circular opening at the top of the heater. The heater can expose samples to a maximum irradiance of approximately 100 kW/m2. Figure  is a schematic of the cone calorimeter.

For piloted ignition tests, an electric spark ignitor is positioned at the top of vertical samples and over the center of horizontal samples. Samples are typically 100 mm by 100 mm, and they can be wrapped with aluminum foil to minimize edge effects. Combustion products and dilution air are extracted through the hood and exhaust duct by a high temperature fan. The flow rate can be adjusted between 0.01 and 0.03 m3/sec. The volumetric flow rate is kept constant during testing. The sample is mounted on a load cell to determine mass loss rate during a test, and smoke obscuration is measured using a laser light source. The gas flow rate in the exhaust duct is calculated from the pressure drop across and temperature at an orifice plate in the duct. Finally, the concentrations of oxygen, carbon dioxide, carbon monoxide, and other gases are measured using appropriate instruments. Heat release rate is calculated from the gas concentration and mass flow measurements.

For this study samples were cut from painted gypsum wallboard panels that were prepared with the panels used in the compartment experiments. Key measurements made with the cone calorimeter, in addition to heat release rate per unit area of the samples, included using a series of cone calorimeter experiments at different heat fluxes to determine a ignition time and surface temperature pairs.

@@

Normally the radiant panel test is used as documented in ASTM~E162 and provides a measure of downward, opposed-flow flame spread..63  The apparatus measures the surface flammability of materials using a gas-fired radiant panel that is 305~mm × 457~mm. However for the purposes of this study a modified version of the test apparatus was used.  The radiant panel and and the radiant panel controls remained in place as documented in ASTM~E162. However the sample mounting system was changed. A rail and trolley system was developed to expose fire fighter protective equipment to various heat fluxes by being able to position the test specimen to a pre-determined position that correlated with a given heat flux.  This trolley was used to hold a sample of painted gypsum wallboard were exposed to the radiant .   Figure

The exposed surface of the samples was parallel to the radiant panel.  The radiant panel was calibrated and mapped with a heat flux gauge. With the flux gauge data, positions away from the radiant panel were identified to enable the samples to be exposed to a range of heat fluxes.  After a timed exposure the samples were examined and amount of thermal damage to the painted paper surface of the gypsum board was   A relative index of flame spread is calculated based on the distance the sample burns. This test method is referenced for building products as ASTM E162, Standard Test Method for Surface Flammability of Materials using a Radiant Energy Source64 and for cellular plastics as ASTM D3675, Standard Test Method for Surface Burning Characteristics of Building Materials.65 Figure 2.3.11 illustrates the radiant panel test apparatus.

@@

>>>>

The cone calorimeter is one of the most commonly used bench-scale rate of heat release apparatus that uses the oxygen consumption method. It was developed at NIST in the 1980s~\cite{babrauskas:1984}. The cone calorimeter has been adopted as ASTM E1354, Test Method for Heat and Visible Smoke Release Rates for Materials and Products Using an Oxygen Consumption Calorimeter~\cite{ASTM E1354}. The cone calorimeter consists of a heater, spark ignitor, sample holder, and load cell located underneath an exhaust hood. Typically, the sample is located in the open with free access of air to the combustion zone. The heater consists of a 5~kW electrical heating element inside an insulated stainless-steel conical shell. Samples can be tested in either a horizontal or vertical orientation. When tests are performed in the horizontal configuration, the specimen is positioned approximately 25~mm beneath the bottom plate of the cone heater. Flames and products of combustion pass through a circular opening at the top of the heater. The heater can expose samples to a maximum irradiance of approximately 100~kW/m\$\^2\$. Figure  is a schematic of the cone calorimeter.

For piloted ignition tests, an electric spark ignitor is positioned at the top of vertical samples and over the center of horizontal samples. Samples are typically 100~mm squares. Combustion products and dilution air are extracted through the hood and exhaust duct by a high temperature fan. The flow rate can be adjusted between 0.01 and 0.03 m\$\^3\$/s. The volumetric flow rate is kept constant during testing. The sample is mounted on a load cell to determine mass loss rate during a test, and smoke obscuration is measured using a laser light source. The gas flow rate in the exhaust duct is calculated from the pressure drop across and temperature at an orifice plate in the duct. Finally, the concentrations of oxygen, carbon dioxide, carbon monoxide, were measured using appropriate instruments. Heat release rate is calculated from the gas concentration and mass flow measurements.

For this study samples were cut from painted gypsum wallboard panels that were prepared with the panels used in the compartment experiments. The experiments were conducted with the sample in the horizontal position.  Measurements made with the cone calorimeter, in addition to heat release rate per unit area of the samples, included using a series of experiments at different heat fluxes to determine a ignition time and surface temperature pairs.

@@

The radiant flooring panel test method is known as ASTM~E648, Standard Method of Test for Critical Radiant Flux of Floor-Covering Systems Using a Radiant Heat Energy Source~\cite{ASTM_E648}. A 1~m long test specimen is mounted horizontally beneath a pre-mixed air-gas-fueled radiant panel. The panel is inclined at 30~\$\^°\$, providing a heat flux on the test specimen varying from 10~kW/m\$\^2\$ at the point closest to the panel to 1~kW/m\$\^2\$ at the farthest point. A pilot flame is used to ignite the test sample. The objective of the test is to determine the critical radiant flux for a given sample.  The critical radiant flux corresponds to the flux at the point of maximum flame propagation, or the point on the sample where the flame spread stopped. For this study the samples examined were cut from painted gypsum wallboard panels that were prepared with the panels used in the compartment experiments.

@@

The vertical radiant panel is used as documented in ASTM~E162 and normally provides a measure of downward, opposed-flow flame spread~\cite{ASTME162}.  The apparatus measures the surface flammability of materials using a pre-mixed air-gas-fired radiant panel that is 305~mm × 457~mm.

However for the purposes of this study a modified version of the test apparatus was used.  The radiant panel and and the radiant panel controls remained in place as documented in ASTM~E162. However the sample mounting system was changed. A rail and trolley system was developed by Lawson and Twilley to expose fire fighter protective equipment to various heat fluxes by being able to move the test specimen to a predetermined position that correlated with a given heat flux.  This trolley was used to hold a sample of painted gypsum wallboard were exposed to the radiant .   Figure

In the modified test configuration the exposed surface of the sample was parallel to the radiant panel. The radiant panel was calibrated and mapped with a heat flux gauge. With the flux gauge data, positions away from the radiant panel were identified to enable the samples to be exposed to a range of heat fluxes.  After a timed exposure the samples were examined and area of thermal damage to the painted paper surface of the gypsum board was compared with heat flux mapping.  This was another means of potentially obtaining a measure of the critical heat flux. Figure 2.3.11 illustrates the radiant panel test apparatus.
